# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

Load the data from the zip file:
```{r}
unzip("activity.zip")
data <- read.csv("activity.csv")
```

Use the date format for the date column:
```{r}
data$date <- as.Date(data$date, format="%Y-%m-%d")
str(data)
head(data)
```

Here is some stats on the dataset:
```{r}
sum(is.na(data$steps))/length(data$steps)
```

## What is mean total number of steps taken per day?

Let's plot the total number of steps by day:
```{r}
library(ggplot2)
library(plyr)
meanByDay <- ddply(data, .(date), summarize, sum = sum(steps, na.rm=TRUE), mean = mean(steps, na.rm=TRUE), median = median(steps, na.rm=TRUE))
ggplot(data=meanByDay, aes(x=date)) + geom_histogram(aes(y=sum), stat="identity", fill = "blue")
ggplot(data=meanByDay, aes(x=date)) + geom_histogram(aes(y=mean), stat="identity", fill = "blue")
ggplot(data=meanByDay, aes(x=date)) + geom_histogram(aes(y=median), stat="identity", fill = "blue")

```

And then the median and the mean for the whole dataset:
```{r}
meanData <- mean(data$steps, na.rm=TRUE)
medianData <- median(data$steps, na.rm=TRUE)
```

The mean of steps is `r meanData`, and the median is `r medianData`.

## What is the average daily activity pattern?

```{r}
sumByInterval <- ddply(data, .(interval), summarize, mean=mean(steps, na.rm=TRUE))
intervalWithMaxSteps <- sumByInterval$interval[which.max(sumByInterval$mean)]

ggplot(sumByInterval, aes(x=interval, y=mean, group=1)) + geom_line() + geom_vline(xintercept=intervalWithMaxSteps, col="Blue")
```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
